---
title: "Prosjektoppgave"
author: "31"
format: html
editor: visual
---

```{r}
rm(list=ls())
```

```{r}
# Packages for data wrangling. 
suppressPackageStartupMessages({
library(haven)
library(tidyverse)
library(lubridate)})
```

```{r}
category <- c("Analgesics","Bath Soap","Beer","Bottled Juices","Cereals",
              "Cheeses","Cigarettes","Cookies","Crackers","Canned Soup",
              "Dish Detergent","Front-end-candies","Frozen Dinners","Frozen Entrees",
              "Frozen Juices","Fabric Softeners","Grooming Products","Laundry Detergents",
              "Oatmeal","Paper Towels","Soft Drinks","Shampoos","Snack Crackers",
              "Soaps","Toothbrushes","Canned Tuna","Toothpastes","Bathroom Tissues")

letter2number <- function(x) {utf8ToInt(x) - utf8ToInt("A") + 1L}
seed_number <- sum(letter2number("Gustav"))
set.seed(seed_number)
sample(category, 1)

```

In this project assignment for the class 1005 - datavitenskap for økonomer, i have used data from the James M. Kilts Center, University of Chicago Booth School of Business. The task uses four datasets. The coustomer count file, and the store-level demographic file. These files can be downloaded from: <https://www.chicagobooth.edu/research/kilts/datasets/dominicks>. The other two files used contains data on laundry detergents (UPCs and Movements). These files can downloaded directly from the links below.

UPC file:

<https://www.chicagobooth.edu/-/media/enterprise/centers/kilts/datasets/dominicks-dataset/upc_csv-files/upclnd.csv?la=en&hash=99A99F1AF79BB754643A1610C67ABA0E4AFAFAB5>

Movement file:

<https://www.chicagobooth.edu/-/media/enterprise/centers/kilts/datasets/dominicks-dataset/movement_csv-files/wlnd.zip?la=en&hash=BFC61F369725363F7F8360CA7D34C1731C985BA9>

Once the files are downloaded locally, u can read them by insterting the file paths in the code block under.

```{r}
ccount <- read_dta('ccount.dta') # Customer count file. 
demo <- read_dta('demo.dta') # Store demographic file. 

# Laundry detergent upc file. 
upc_ld = read.csv('upclnd.csv',sep=',') 

# Laundry detergent movement file. 
mvm_ld = read.csv('wlnd',sep=',')
```

```{r}
# Run this for different detergent UPCs:

# unique(upc_ld$DESCRIP)
```

We can see there is a lot of different UPCs for detergents. I want to choose some brands for the data analysis. Beacuse the data is so big, i am only gonna pick three brands.

```{r}
brands <- paste(c("SURF", "WISK", "WOOLITE"), collapse = "|") # Picking brands for data analysis. 

upc_ld <- upc_ld %>%
  filter(grepl(brands,DESCRIP)) # Filtering brands. 

upc_ld <- upc_ld %>%
  mutate(BRAND = str_extract(DESCRIP, brands)) # Adding brands as column. 
```

```{r}
#|message: false
mvmupc <- left_join(upc_ld,mvm_ld, by=c('UPC')) # Left join by key. 

# Adding sales column in dollars. 
mvmupc <- mvmupc %>%
  group_by(WEEK,STORE,UPC) %>% # Group by keys. 
  mutate(SALES = PRICE*MOVE/QTY) %>% # Formula as stated in data manual. 
  filter(SALES > 0) # Filter for sales over 0. 

mvmupc <- mvmupc %>% # Relocateing columns for tidiness.
  relocate(SALES, .before = PROFIT) %>% 
  relocate(WEEK, .before = COM_CODE)  %>% 
  relocate(STORE, .before = COM_CODE)

mvmupc <- mvmupc %>%  # Filter for weeks in 1990
  filter(between(WEEK,16,68))

mvmupc <- mvmupc %>% # Removing some variables that i dont need.
  select(-COM_CODE,-PRICE_HEX,-PROFIT_HEX,-NITEM)  

mvmupc <- mvmupc %>% # Calculating some data i need for analysis. 
  group_by(STORE, WEEK, BRAND) %>%
  summarise(MOVE_SUM = sum(MOVE),
            AVG_PROFIT = mean(PRICE),
            SUM_PROFIT = sum(PROFIT),
            AVG_PRICE = mean(PRICE),
            SUM_SALES = sum(SALES))
```

```{r}
colnames(ccount)
```

```{r}
ccount <- na.omit(ccount) # Removing missing data. 

ccount$date <- as.Date(ccount$date, "%y%m%d") # Date variable.  

ccount <- ccount %>% 
  mutate(date = as.Date(date)) %>% # Convert to date. 
  filter(between(date, as.Date('1990-01-01'), as.Date('1990-12-31'))) %>% # Filter for year. 
  relocate("week", .before = store) %>% # Relocate for tidiness. 
  rename(WEEK = week) %>% # Renaming for merge. 
  rename(STORE = store)
```

```{r}
# Using the colnames() function to get var names. 

# Aggregate by week and store
ccount <- ccount %>%   
  group_by(WEEK,STORE) %>% 
  summarise_at(.vars = c("grocery","dairy","frozen","bottle","mvpclub","groccoup","meat","meatfroz","meatcoup","fish","fishcoup", "promo","promcoup", "produce",  "bulk","saladbar","prodcoup","bulkcoup","salcoup","floral","florcoup", "deli", "deliself", "deliexpr","convfood", "cheese","delicoup","bakery" , "pharmacy", "pharcoup", "gm" , "jewelry", "cosmetic","haba","gmcoup",   "camera" ,  "photofin" ,"video"  , "videoren", "vidcoup" , "beer" , "wine" ,"spirits" , "miscscp" , "mancoup" , "custcoun", "ftgchin" , "ftgccoup","ftgital" , "ftgicoup" ,"daircoup" ,"frozcoup", "habacoup" ,"photcoup" ,"cosmcoup", "ssdelicp" ,"bakcoup" , "liqcoup"), .funs = sum) 
```

For a better overview i have printed the number of store observations in both the merged movementfile and the aggregated ccount file.

```{r}
unique(mvmupc$STORE)
```

```{r}
unique(ccount$STORE)
```

We can now see that the ccount includes more data on stores. If we merge the data there will be some missing values on store observations.

```{r}
ccount_mvmupc <- left_join(mvmupc, ccount, by= c("WEEK","STORE"))

ccount_mvmupc <- ccount_mvmupc %>% # Removing duplicated rows (if any).  
  distinct()
```

```{r}
demo <- demo %>% # Removing column with missing values. 
  select(-gini)

demo <- demo %>%  # Renaming for merge. 
  rename(STORE = store) 

sum(is.na(demo))

# We can see there is a lot of missing values in the demo data. We have to keep that in mind for later analysis. 
```

```{r}
df <- left_join(ccount_mvmupc,demo,by=c('STORE')) # Left join by key.

# Weeks start at 16, to make variable start at 1 i substract by 15.
df <- df %>% 
  mutate(WEEK = WEEK - 15)

df <- df %>% 
  distinct() # Removing duplicated rows (if any). 

df <- df[, 1:55]

df <- df %>%
  select(-ftgchin,-pharmacy) # Removing column that contains 0 observations (Food to go Chinese) and pharmacy.
```

```{r}
write_csv(df, "df.csv") # Writing as csv file. 
```
